---
title: "Presentation"
author: "Names"
format: 
  revealjs:
    embed-resources: true
    theme: "default" 
---

# Intro to Slides with Quarto (Delete This Section for Final Presentation!)

## New Slides {.smaller}

- `#` in the text editor creates a *section* header slide, with text in the 
centered in the middle of the slide
- `##` in the text editor creates a *new slide*: the text next to the `##` 
is the title of the slide
- Note also (which you can only see in the `.qmd`) that the `{.smaller}`
code next to a slide title makes the text of that slide smaller (compare the
text size of this slide to the text size in the next slide)

## Bulleted Lists 

- You can use `-` in the text editor to make bulleted lists
- If you want to have bullet points appear click-by-click, you can 
use the **incremental** div:
```
:::{.incremental}
- these bullet points
- will appear click by click
:::
```

## Slide with Columns 

::::{.columns}

:::{.column width=50%}
Column 1
:::

:::{.column width=50%}
Column 2
:::


::::

## Bulleted Lists Click-By-Click Example

:::{.incremental}
- these bullet points
- will appear click-by-click
:::

## Embedding R Code

- Embedding R code works just like in regular Quarto, using R chunks

```{r}
#| echo: true
mtcars
```

## Visualization

```{r}
#| echo: true
library(tidyverse)
ggplot(mtcars, aes(x = mpg, y = wt)) +
  geom_point()
```

## Customization

- You can change the default aesthetic of your slide show by 
changing the *theme* by using the `theme` YAML option (you have to add it 
in the `revealjs` section, beneath `embed-resources`)
- See [here](https://quarto.org/docs/presentations/revealjs/themes.html) for 
the themes you can use (and adapt - you can also make your own if
you want, see the instructions in the linked page)
- Theme names are put in "quotes"

## Resources

- You find Quarto revealjs documentation [here](https://quarto.org/docs/presentations/revealjs/)
- Meghan Hall also has a [page](https://meghan.rbind.io/blog/quarto-slides/) 
with useful tips

# Presentation Template

## Note

- You do not have to use this format exactly!
- Presentations should be roughly 7 minutes
- Each group member should speak for an equal amount of time
- You can separate sections with section slides but do not have to

## Introduction to Topic and Motivation

## Introduction to Data

## Highlights from EDA 1

## Highlights from EDA 2

## Inference and Modeling 1

## Inference and Modeling 2

## Conclusions and Future Work
