<<<<<<< HEAD
```{r process the data}
library(vdemdata)
a=load("/cloud/project/data/vdemdata-master/data/vdem.RData")
summary(a)

```
=======
---
title: "Data Exploration"
author: "Group 4"
date: "7/6/2023"
format: pdf
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

# Project Proposal 

## Research Question
Our research question idea explores the relationship  between prosocial/
social behaviors and civil society index. This question is of interest because
it pulls together politics and measures of wellbeing to explore what effects 
civil society, which is the area outside of business and government, such as 
family and community. We want to explore how various measures of community 
engagement and support, both positive and negative, predict measure civil
society scores. This will allow us to practice both clustering and regression 
modeling to answer this question. We will also be able to practice cleaning and 
merging data. While we cannot say for certain what relationships we expect to 
see, finding relationships between the democracy and happiness index datasets is
an exciting possibility. Specific relationships we could see include a 
potentially negative between civil society organization oppression, freedom to 
make life choices, and wars/coups, a potentially positive relationship between 
social support, participation  rate in civil society, generosity, education, 
and maybe government corruption. We expect places with high civil society 
indexes to have higher amounts of things that bring people together and lower 
rates of things that bring people apart or make meeting with other people 
difficult. 

## Data
Since our topic is about how civil society index be affected by prosocial 
behaviors, we will use the variables listed above, which include: civil society
repression effort scores, social support scores, freedom to make life choices, 
presence of war, presence of coups, participation rate in civil society, generosity, 
education, and government corruption. The data sets we will be using are vdem 
(democracy) and wellness. Combining these variables allow us to look at 
how civil society and connections between people exist from different lenses. 
Combining data that is political and sociological in nature allows us to get a fuller
perspectice and understanding of the relationship between different behaviors that 
connect people, prosocial behaviors, and their relationship to civil society in 
different countries. 

In the first part of the data processing, we will determine which of the factors
will be important/influentional/significant in predicting civil society index (or participation...). 
After choosing the significant factors, both supervised and 
unsupervised methods could be conducted by either linear models and clustering,
or if we focus on prediction/grouping, doing k-means clustering and hierarchical 
clustering. This would allow us to explore the difference in k-means and 
hierarchical clustering to see how the prediction/results change. 

# Exploring the Data
Answer the following questions:

What is your outcome variable(s)? How well does it measure the outcome you are 
interested in? How does it relate to your hypotheses? What are your key 
explanatory variables? In addition, create a table of summary statistics for 
the variables you are planning to use.

## Merging data
```{r load-data}

load("/cloud/project/data/vdemdata-master/data/vdem.RData")

```

## Cleaning data 
- Count how many NA observations we have. If it is under %, we will omit the obs.
or we could assign some values to them using given method.
- Deal with country name changes 

Our dependent variable is civil society participation. We will use the predictor 
variables listed above, which include: civil society repression effort scores, 
social support scores, freedom to make life choices, presence of war, 
presence of coups, participation rate in civil society, generosity, education, 
and government corruption. 




>>>>>>> 42f66bc75e801845634806ecf30a37ad052fc256
