---
title: "Data Exploration"
author: "Group 4"
date: "7/6/2023"
format: pdf
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

# Exploring the Data
Answer the following questions:

What is your outcome variable(s)? How well does it measure the outcome you are 
interested in? How does it relate to your hypotheses? What are your key 
explanatory variables? In addition, create a table of summary statistics for 
the variables you are planning to use.

General Answer: Our dependent variable is 
civil society participation. v2x_cspart

We will use the predictor variables listed above, which include: 
civil society repression effort scores,  v2csreprss
social support scores,  happiness
freedom to make life choices,  happiness
presence of war, e_civil_war
presence of coups, e_pt_coup
participation rate in civil society, v2x_partipdem
generosity, happiness
education, e_peaveduc
and government corruption. v2x_corr

## Merging data - James
```{r load-data}
library(readxl)
wh_2023 <- read_excel("/cloud/project/data/wh_2023.xls")
#View(wh_2023)
load("/cloud/project/data/vdemdata-master/data/vdem.RData")
# Merge data by country and year/time
colnames(wh_2023)
colnames(vdem)
```

```{r data processing(merge)}
need_vdem=c("country_name","year","v2csreprss",'e_civil_war',"e_pt_coup","v2x_partipdem","e_peaveduc","v2x_corr","v2x_cspart")
vdem_use=vdem[,need_vdem]
need_wh=c("Country name","year","Social support","Freedom to make life choices","Generosity")
wh_use=wh_2023[,need_wh]
colnames(wh_use)[1] <- "country_name"
total=merge(vdem_use, wh_use, by = c("country_name", "year"),all.x=TRUE,all.y=TRUE) 
```

## Cleaning data - Alicia 
- Count how many NA observations we have. If it is under %, we will omit the obs.
or we could assign some values to them using given method.
- Deal with country name changes - optional
- Drop obs or columns not needed 

- Histograms and box plots 
- Create a table of summary statistics for the variables you are planning to use.

## Code book - README.qmd - Kaori
- Location must be specified

## Data Visualization - Ephrata and Kaori 
Question 1:  
How well does it measure the outcome you are interested in? - Kaori

Question 2: 
What is our hypothesis? - Kaori 
How do our variable choices relate to our hypotheses? (paraphrasing from the proposal) - Kaori 

Visualization group 1: 
- A correlation heat map - Kaori 

<<<<<<< HEAD
>>>>>>> 42f66bc75e801845634806ecf30a37ad052fc256
=======
Visualization group 2: 
- Scatter plots civil society participation vs. each single predictor - Ephrata
- Line graphs showing changes over time for each predictor vs. outcome - Ephrata
- QQ Plot - Ephrata
>>>>>>> 9973f8fb6b863809cff280ca64f138b9de20584f
