# Analysis Experimentation

## Methods Overview
Choose 2 methods that we have talked about and use them to model the outcome 
you are interested in. Present the mathematical intuition behind the method 
and define the model in the context of your data and your research question.
If the models require hyperparameters (tuning parameters) discuss how you plan 
to go about tuning the model.

1) Lasso Regression: We are using lasso regression to choose the best predictors. 
This will allow our clustering method to be more effective by only using the most
influential predictors. Our tuning parameters will be chosen by using cross validation
on the (insert tuning parameters here). 

2) K-means clustering: This will allow us to see the relationships between clusters
of countries with similar characteristics. Being able to understand the cluster 
centers for each variable will provide us with information regarding what is important
in each cluster, highlighting what the key differences between them are. 

## Application
Fit the two models you have presented in the previous section. If this process
involves tuning parameters, include the process of tuning the model.
```{r lasso}


```

```{r k-means}


```


## Visualizing
Depending on the direction of your project produce at least one visualization that

predictive questions: compares model fit between the two methods
inferential questions: compares the way in which the two models help you
understand the relationship between your outcome and your explanatory variable of 
interest.

## Discussion
In one to two paragraphs, discuss what you have learned from your initial modeling attempts.